---
title: "AnalyseGEM"
author: "edmar.teixeira@me.com"
date: "Sunday, December 07, 2014"
output: html_document
---

# Background

## Aim

Quantify **uncertanty** (sensitivity of model to?) coming from the choice of (possible?) **sowing dates and maize hybrids** in regional simulations

## Method

> Use APSIM 7.4

> Run model for ~2700 pixels of 5 x 5 km (NIWA ref XXX) across New Zealand

> Arable land filter frpm LandCare (ref XXX)

> Run 30 years (XXX-XXXX) of data from NIWA (XXXX)

> Run 5 hybrid maturity types (parameter XXXX with values x, x, x, x, x)

> Run at 5 different sowng dates (1-Sep, 1-Oct, 1-Nov, 1-Dec, 1-Jan)


# Data Manipulation

```{r, echo=TRUE, include=FALSE}
library(ggplot2)
library(dplyr)
```

## Retrieve data and check
```{r, include=FALSE}
# load data from HD
#setwd("E:\\SowByGenotype")
setwd("C:\\Apsim_dev\\Projects\\2014-SowByGenotype\\GEM_data")

load("DATA_GEM.Rda", .GlobalEnv)


head(DATA_GEM)
tail(DATA_GEM)
summary(DATA_GEM)
```

> Subset to keep only maize (exclude winter forage wheat) and order factors
```{r, include=FALSE}
# Remove wheat (catch crop rotation) and keep only maize data

workDF = filter(DATA_GEM, CurrentCrop != 'wheat_exceed')

# order sow dates
workDF$thisSow = factor(workDF$thisSow, levels = c('s1', 's2', 's3', 's4', 's5'))
```

## Graph some results

> Total silage biomass by sowing date (kg/ha)
```{r}

# Pooled response to sow date
ggplot(workDF, aes(x= thisSow , y=TotalBiomass)) +  geom_boxplot() +
facet_grid( .~ thisHyb)

```

> Total silage biomass by hybrid type (kg/ha)
```{r}

# Pooled response to Hybrid (Bio)
ggplot(workDF, aes(x= thisHyb , y=TotalBiomass)) +  geom_boxplot() +
facet_grid( .~ thisSow)

```

> Total **grain yield** by hybrid type (kg/ha)
```{r}

# Pooled response to Hybrid (Bio)
ggplot(workDF, aes(x= thisHyb , y=TotalYield)) +  geom_boxplot() +
facet_grid( .~ thisSow)

```

> Harvest index by hybrid type
```{r}

# Pooled response to Hybrid (HI)
ggplot(workDF, aes(x= thisHyb , y=HarvestIndex)) +  geom_boxplot() +
facet_grid( .~ thisSow)

```

# Analysis

## Calculate variability 

> Just to have an initial idea of variability

> Considers variation among YEARS and PIXELS together

> Pooled among hybrids and sowing dates


```{r}
# polled CV across hybrids and sowing dates
cvTable = workDF %>%
  group_by(thisSow, thisHyb) %>%
  summarise(CVperc = sd(TotalBiomass, na.rm = TRUE)/mean(TotalBiomass, na.rm = TRUE)*100)

ggplot(cvTable, aes(x= thisHyb , y=CVperc)) +  geom_point() +
facet_grid( .~ thisSow)

```

## Values pooled (median, mean, stdev) across **years** 

> Statistics of n years per pixel/scenario of hybrid and sow date combination

> Also creates a ID for each pixel-h-s combination (FIXME: maybe redundant now)

```{r}
# Create a median and stdev for each pixel-h-s, polled across YEARS (i.e. year is not included in grouping)

workDF_med = workDF %>%
  group_by(row,col,thisHyb,thisSow) %>%
  summarise( thisLat = mean(thisLat, na.rm = TRUE),
             thisLong = mean(thisLong, na.rm = TRUE),
             Bio_mean = mean(TotalBiomass, na.rm = TRUE),
             Bio_med = median(TotalBiomass, na.rm = TRUE),
             Bio_sdv = sd(TotalBiomass, na.rm = TRUE),
             Bio_cv = (Bio_sdv/Bio_mean)*100,
             Y_mean = mean(TotalYield, na.rm = TRUE),
             Y_med = median(TotalYield, na.rm = TRUE),
             Y_sdv = sd(TotalYield, na.rm = TRUE),
             Y_cv = (Y_sdv/Y_mean)*100,
             HI_med = median(HarvestIndex, na.rm = TRUE)) %>%
  # Unique ID for the pixel in a hxs scenario (FIXME: maybe redundant now)
  mutate(pixel_scenID=paste0(row,"_",col,"_",thisHyb,"_",thisSow),
         pixel_ID=paste0(row,"_",col))   

# make hybrid and sow date as numbers (for using in raster later) - FIXME: do this prettier later
workDF_med$thisSow_n[workDF_med$thisSow == "s1"] <- 1
workDF_med$thisSow_n[workDF_med$thisSow == "s2"] <- 2
workDF_med$thisSow_n[workDF_med$thisSow == "s3"] <- 3
workDF_med$thisSow_n[workDF_med$thisSow == "s4"] <- 4
workDF_med$thisSow_n[workDF_med$thisSow == "s5"] <- 5

workDF_med$thisHyb_n[workDF_med$thisHyb == "h1"] <- 1
workDF_med$thisHyb_n[workDF_med$thisHyb == "h2"] <- 2
workDF_med$thisHyb_n[workDF_med$thisHyb == "h3"] <- 3
workDF_med$thisHyb_n[workDF_med$thisHyb == "h4"] <- 4
workDF_med$thisHyb_n[workDF_med$thisHyb == "h5"] <- 5

head(workDF_med)
tail(workDF_med)

```


Select some pixels across the range of results
```{r}
library(dplyr)
# Give me the 25th percentile of Biomass
df_p <- workDF_med %>% group_by(row, col) %>%
  mutate(perc = rank(Bio_med)/length(Bio_med))

# select pixels to graph

df_sel = data.frame(rowsSel = c(67, 184, 239), colsSel = c(180, 113, 38) )


for (p in 1:nrow(df_pix)) {
  
    

df_pix <- df_p %>% group_by(row, col) %>%
  #filter(row==184 & col==113)
  filter(row==df_sel$rowsSel[p] & col==df_sel$colsSel[p])

head(df_pix, 25)
write.table(df_pix, file=paste0(df_sel$rowsSel[p],"_",df_sel$colsSel[p],"_","3D_graph.txt"), row.names=FALSE, col.names=TRUE, sep=" ",
               append=FALSE, quote=FALSE)

par(mfrow = c(1, 2))
# PLot 3 D graph
z = matrix(df_pix$Bio_med, # the data elements 
nrow=length(unique(df_pix$thisHyb)),              # number of rows 
ncol=length(unique(df_pix$thisSow)),              # number of columns 
byrow = TRUE)        # fill matrix by rows 
colnames(z) = unique(df_pix$thisHyb_n)
rownames(z) = unique(df_pix$thisSow_n)

x = as.numeric(colnames(z))
y = as.numeric(rownames(z))
  
persp3D(x, y, z, theta=60, phi=10, expand=0.75, col=color[max(z)],
        ticktype="detailed", lighting = TRUE, xlab="Sow date", ylab="Hybrid", zlab="Biomass (t/ha)",axes=TRUE)


# variaility
# PLot 3 D graph
z = matrix(df_pix$Bio_sd, # the data elements 
nrow=length(unique(df_pix$thisHyb)),              # number of rows 
ncol=length(unique(df_pix$thisSow)),              # number of columns 
byrow = TRUE)        # fill matrix by rows 
colnames(z) = unique(df_pix$thisHyb_n)
rownames(z) = unique(df_pix$thisSow_n)

x = as.numeric(colnames(z))
y = as.numeric(rownames(z))
  
persp3D(x, y, z, theta=60, phi=10, expand=0.75, col=color[max(z)],main= paste0(df_sel$rowsSel[p],"_",df_sel$colsSel[p]),
        ticktype="detailed", lighting = TRUE, xlab="Sow date", ylab="Hybrid", zlab="Biomass (t/ha)",axes=TRUE)

}

```




## Find hybrid x sow date with **highest yield** in each pixel

> This is to check the data "sensibility"

> FIXME: Check this dplyr script and logic as it seems that almost  **NO variation** is coming from this 

> Most productive sowing dates are **S1 (Sept) and S2 (Oct) **

> Most productive is always **hybrid 5** (longest hybrid)


```{r, echo=FALSE}
# select line with maximum yield per pixel
df_maxB <- workDF_med %>% group_by(row, col) %>%
  filter(Bio_med == max(Bio_med))

head(df_maxB)
nrow(df_maxB)

#summary(df_maxB)

#PLot histograms
hg1 = ggplot(df_maxB, aes(x=thisSow_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")
hg2 =ggplot(df_maxB, aes(x=thisHyb_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")

# Plot
require(gridExtra)
grid.arrange(hg1, hg2,ncol=2)
```


## Find highest **grain yield** per pixel

> Allmost same results for grain

```{r, echo=FALSE}
# select line with maximum yield per pixel
df_maxY <- workDF_med %>% group_by(row, col) %>%
  filter(Y_med == max(Y_med))

head(df_maxY, 25)
nrow(df_maxY)

#summary(df_maxY)

#PLot histograms
hg1 = ggplot(df_maxY, aes(x=thisSow_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")
hg2 = ggplot(df_maxY, aes(x=thisHyb_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")


# Plot
require(gridExtra)
grid.arrange(hg1, hg2,ncol=2)
```


## Find hybrid x sow date with **lowest inter-annual variability** in each pixel

> A bit more spread of frequencies here

> In most pixels, sowing date 2 and 3 are less variable

> Longer hybrids (Hybrid 3,4 and 5) are less variable


```{r}
df_minSd <- workDF_med %>% group_by(row, col) %>%
#  filter(Bio_sdv == min(Bio_sdv))
 filter(Bio_cv == min(Bio_cv))

head(df_minSd)
nrow(df_minSd)
#summary(df_minSd)

#summary(df_maxY)

#PLot histograms
hg1 = ggplot(df_minSd, aes(x=thisSow_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")
hg2 =ggplot(df_minSd, aes(x=thisHyb_n)) + geom_histogram(binwidth=.5, colour="black", fill="white")


# Plot distributions of Si
require(gridExtra)
grid.arrange(hg1, hg2, ncol=2)

```

# How is total biomass and inter-annual variability related?

> In general **highest** yields have **lowest variability**

```{r, echo=FALSE}

workDF_med$Bio_meanTons = workDF_med$Bio_mean / 1000 # convert kg/ha to t/ha

# all together
ggplot(workDF_med, aes(x= Bio_meanTons , y=Bio_cv)) +  geom_point(aes(colour = factor(thisSow)) ) +
facet_grid(thisHyb ~.)

#different panels by hybrid
ggplot(workDF_med, aes(x= Bio_meanTons , y=Bio_cv)) +  geom_point(aes(colour = factor(thisHyb )) ) +
facet_grid(thisSow ~thisHyb, scales="free")

```

# Rasterise for mapping

> 

> Checking data-frame

> Plot statistics for n Years 

## Pixels with highest Total Silage Biomass

> Maps show stats and which hybrid and sowing date was selected in each pixel

```{r, echo=FALSE, warning=FALSE}
library(raster)
par(mfrow=c(1,1))
outNames = colnames(df_maxY)
outNames =gsub(" ","",outNames )

# ignore first 6 factors
spg = list() # dataframe list to be coerced as spatial object
rast = list() # raster
s=list()
for(o in 7:length(outNames)){
if (o == 12) {next} # other label to ignore

  # For aggregate
  spg[[o]] = data.frame(df_maxY$thisLong, df_maxY$thisLat, df_maxY[[o]])
  coordinates(spg[[o]]) = ~ df_maxY.thisLong + df_maxY.thisLat # Attention to variable names
 
  # Common to both methods (grid, raterise and project)
  gridded(spg[[o]]) <- TRUE
  rast[[o]] = raster(spg[[o]])
  proj4string(rast[[o]]) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

 plot(rast[[o]], main = paste0(outNames[o]))

  # stack them
  if(o == 1) {
    s = stack(rast[[o]])} 
  else {
    s[[o]] = rast[[o]]
  }
  
  # save raster as
  #writeRaster(rast[[o]], file=paste0(cropNames[c],"\\",thisCrop,"_", outNames[o],"_var",".tiff"), overwrite=TRUE) 
}

```


## Show maps when selecting run with **lowest variability**


```{r, echo=FALSE, warning=FALSE}
library(raster)
par(mfrow=c(1,1))
outNames = colnames(df_minSd)
outNames =gsub(" ","",outNames )

# ignore first 6 factors
spg = list() # dataframe list to be coerced as spatial object
rast = list() # raster
s=list()
for(o in 7:length(outNames)){
if (o == 12) {next} # other label to ignore

  # For aggregate
  spg[[o]] = data.frame(df_minSd$thisLong, df_minSd$thisLat, df_minSd[[o]])
  coordinates(spg[[o]]) = ~ df_minSd.thisLong + df_minSd.thisLat # Attention to variable names
 
  # Common to both methods (grid, raterise and project)
  gridded(spg[[o]]) <- TRUE
  rast[[o]] = raster(spg[[o]])
  proj4string(rast[[o]]) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

 plot(rast[[o]], main = paste0(outNames[o]))

  # stack them
  if(o == 1) {
    s = stack(rast[[o]])} 
  else {
    s[[o]] = rast[[o]]
  }
  
  # save raster as
  #writeRaster(rast[[o]], file=paste0(cropNames[c],"\\",thisCrop,"_", outNames[o],"_var",".tiff"), overwrite=TRUE) 
}

```


## Calculate the relative variability for factors

> How much of overall variation is coming from sowing date, hybrid and their interaction?

> Adapting Esther Meenken method of sensitivity analysis for that

> Run a ANOVA and retrive the Sums of Squares (SS) of each factor (2 main factors and the interaction)

> Display it as a relative proportion of total variation (% of total SS)

> This is the **Si index** summary:

```{r, echo =FALSE}
# do an anova for each pixel 
si_df_final = NULL
pixels = unique(workDF_med$pixel_ID)
for (p in 1:length(pixels)) {
  
  thisPixDF = filter(workDF_med, pixel_ID == pixels[[p]])
  
  summary(thisPixDF)
         
  # do anova
  anova = aov(Bio_mean ~ thisHyb*thisSow, data = thisPixDF)
  anovaTable = summary(anova)
  
  #Retrive SS and MS values
  anova_SS = anovaTable[[1]]$'Sum Sq'
  anova_MS = anovaTable[[1]]$'Mean Sq' 

  # Binds all SS for each variable together into a new df
  df_ss = data.frame(SS = anova_SS, MS = anova_MS)

  #retrieves row names and load into result df
  namesAnTab = rownames(anovaTable[[1]])
  rownames (df_ss) = gsub(" ","",namesAnTab) # remove spaces

  # Creates a column with "normalised" percent of total variation (SS) 
  # in relation to total SS for each factor
  df_ss$SS_p = round((df_ss$SS/sum(df_ss$SS))*100,3)
  df_ss$MS_p = round((df_ss$MS/sum(df_ss$MS))*100,3) # FIXME: Not sure which to use SS or MS
 
 
  df_si=data.frame(thisLat = thisPixDF$thisLat[1],
                   thisLong = thisPixDF$thisLong[1], 
                   df_ss$SS_p[1], 
                   df_ss$SS_p[2], 
                   df_ss$SS_p[3])
  
  # label Si df
  colnames(df_si)[3] = gsub(" ","",paste0("Si_",namesAnTab[1]))
  colnames(df_si)[4] = gsub(" ","",paste0("Si_",namesAnTab[2]))
  colnames(df_si)[5] = "Si_Interac"
  
  # Appends to final Si df
  if(p==1){ 
    si_df_final = data.frame(df_si)  
    } else {
    si_df_final = rbind(si_df_final, df_si)
    }
  
}

summary(si_df_final)
nrow(si_df_final)
gH =ggplot(si_df_final, aes(x=Si_thisHyb)) + geom_histogram(binwidth=.5, colour="black", fill="white")
gS = ggplot(si_df_final, aes(x=Si_thisSow)) + geom_histogram(binwidth=.5, colour="black", fill="white")
gI = ggplot(si_df_final, aes(x=Si_Interac)) + geom_histogram(binwidth=.5, colour="black", fill="white")

# Plot distributions of Si
require(gridExtra)
grid.arrange(gH, gS, gI,nrow=3)
```


> Histograms show frequency of pixels in each range of Si for each factor and interactions

## Map Si index

> Si is the fraction of total sums of squares attributed to each main factor or interactions

> Let's plot it: 

```{r, echo = FALSE, warning=FALSE}
library(raster)
par(mfrow=c(1,1))
outNames = colnames(si_df_final)
outNames =gsub(" ","",outNames )

# ignore first 6 factors
spg = list() # dataframe list to be coerced as spatial object
rast = list() # raster
s=list()
for(o in 3:length(outNames)){

  # For aggregate
  spg[[o]] = data.frame(si_df_final$thisLong, si_df_final$thisLat, si_df_final[[o]])
  coordinates(spg[[o]]) = ~ si_df_final.thisLong + si_df_final.thisLat # Attention to variable names
 
  # Common to both methods (grid, raterise and project)
  gridded(spg[[o]]) <- TRUE
  rast[[o]] = raster(spg[[o]])
  #rast = raster(spg[[o]])
  proj4string(rast[[o]]) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
  #proj4string(rast) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

 plot(rast[[o]], main = paste0(outNames[o]))

  # stack them
  if(o == 1) {
    s = stack(rast[[o]])} 
  else {
    s[[o]] = rast[[o]]
  }
  
  # save raster as
  writeRaster(rast[[o]], file=paste0("Si_", outNames[o],".tiff"), overwrite=TRUE) 
}

```




