---
title: "netCDFToMet_test"
author: "edmar.teixeira@me.com"
date: "Wednesday February 11, 2015"  "Friday, January 16, 2015"
output: html_document
---

# Purpose: read netCDF climate from NIWA and create .met files for APSIM 
# plus .frost files
# 16 NOV 2014 Using this "RCP" version for any scenario (updated one)  
# 10 Feb 2015 Removed frost calculation to simplify it
# 24 July 2015 Prepared it for ERA dataset for the GEM work in Bonn University
# 2 Mar 2016 changed to new netcdf library to test script

# Memory to clean up and libraries to load  
```{r Libs, echo=FALSE, include=FALSE}
gc()
memory.limit(size=4095)
library("knitr")
#library("ncdf")
library("chron")
library("plyr")
library("dplyr")
library("ncdf4")
```

# Set RCP first

```{r DefineParam, echo=FALSE, include=FALSE}

setwd("F:\\ERA-40_Reanalysis\\")

RCP <- "ERA" #"RCPpast" # RCPpast RCP8.5 ; using 2.6 for ERA (historical)

climDir <- getwd() # actual directory where climate data should be

```

# CO2 per year function (retrieves from RCP data from NIWA (Brett Mullan 2014)

```{r CO2Set, echo=FALSE}
co2PerRCP  <- 
  read.table(paste0("C:\\Apsim_dev\\Projects\\CCII\\CO2_files\\CO2_",RCP,".txt"), sep="\t", header=T)

# Check CO2
co2PerRCP$co2 <- as.double(co2PerRCP$co2)
head(co2PerRCP)
tail(co2PerRCP)
summary(co2PerRCP)
plot(co2PerRCP)
```

# Alternative CO2 calculation option - comment out if not A2

```{r Co2CalcFunc, echo=FALSE, include=FALSE}
#co2Year = function(y) {0.0285*y^2 - 112.09*y + 110541} 
# empirical fit from NIWA's A2 data used in SLMACC-2012
#Test CO2 function
#test_Year = 1976
#co2Year (test_Year)

# Create a co2 for A2 - comment out if doing other scenarios
#co2PerRCP = data.frame(year = c(1971:2120))
#co2PerRCP$co2 = co2Year(co2PerRCP$year)
head(co2PerRCP)
tail(co2PerRCP)
```

# Define file naming of netCDFs

```{r OrderDataLabel}

# ---- Attention (names as per NIWA standard)

factorName <- c('maxt','mint','radn', 'rain') # renamed .nc for simplification

# Obs: rain is last factor to speed the finding of -9999 or NA in maxt records
```

# Define the first year of each netCDF file <---- Change Period

```{r SetFirstYears}

#firstYear <- c(1971,1981,1991,2001,2011,2021,2031,2041,2051,2061,2071,2081,2091) 
#firstYear <- c(1971,1981,1991)
firstYear <- c(1971, 1981,1991)
#firstYear <- c(2081,2091)

# ---- Attention changed name of last netCDF to match name convention

# label for folder that will store .met
runPeriod <- paste0(firstYear[1],"_",max(firstYear)+9)
```

```{r SetOutFolder}
# set output folder
outFolder <- paste0(climDir,"/metFiles_Test_",RCP,"_",runPeriod,"/")
suppressWarnings(dir.create(outFolder))
```

# Data conversion and other date functions 

```{r SetFuncs}
# Functions used for weather data conversion
kToC <- function(k) {return(k-273.15)} # Converts kelvin to celcius (oC)

WattToMJ <- function(w) {return(w*0.0864)} # Converts watts/m2 to MJ/m2

# day of year (1-366)
doy.func <- function (dayCount, startYear) {
  date <- chron(dayCount-1, 
               format = c(dates = "dd/mm/yy"), 
               origin=c(month=1,day=1,year=startYear))
  
  doy <- as.numeric(strftime(date, format = "%j"))  
  
  return (doy)
}

# year function
year.func <- function (dayCount, startYear) {
  date <- chron(dayCount-1, 
               format = c(dates = "dd/mm/yy"), 
               origin=c(month=1,day=1,year=startYear)) 
  year <- as.POSIXlt(date)$year+1900
  return (year)
}
```

# Find and create the temporal dimensions as day of year (doy) and year
```{r GetTime}

fileName <- NULL

files <- list.files(climDir,pattern='.nc', full.names=FALSE) # find files
files
# creates day and year common vectors in APSIM format for whole period
thisFile <- NULL
doy <- NULL  # vector with appended doy
year <- NULL # vector with appended year

# loop years
for (y in 1:length(firstYear)) {
  
  # Get one "sample"" file to get doy and year
  thisFile <- paste(factorName[4],"_",firstYear[y],"-",firstYear[y]+9,".nc", sep = "", collapse = NULL)
  
  # open netCDF
  base.nc <- nc_open(paste0(climDir,"\\",thisFile)) # new lib ncdf4

  
  # loop daily to get doy and year dimensions
  for (t in 1:length(ncvar_get(base.nc, "time"))) {       
    # FIXME: Dealing with bug in NIWAs data, some years should be leep after 2040
    # So 2040 on the doy and year function gine NAs on some years
    # This needs checking in final version
    if(is.na(doy.func(t,firstYear[y]))) { 
      doy <- c(doy, 1) # make it doy 1
      year <- c(year, (1+year.func(t,firstYear[y]))) # make it next year
      print(paste(c("Bug in NIWAs data for doy calculation:",
                    t, y, 
                    year.func(t,firstYear[y]), 
                    doy.func(t,firstYear[y]))))
    } else {
      doy <- c(doy, doy.func(t,firstYear[y]))
      year <- c(year, year.func(t,firstYear[y])) 
    }
    rowNo <- base.nc$dim$latitude$len
    colNo <- base.nc$dim$longitude$len
  }
  nc_close(base.nc)
  
}

# check lengths of vectors are the same
length (doy)
length (year)
```

```{r setStartTime}
timeStart = Sys.time() # just for control of simulation time
```

# Read data

Loop through netCDF files

```{r ReadNetCDFData}

#TODO: Select arable grid-cells

LUFilter <- 
 # read.table("C:\\Apsim_dev\\Projects\\2014-SowByGenotype\\gisLayers\\luFilter\\LandUseFilter_Allcats.txt", 
   read.table("C:\\apsim_dev\\Projects\\CCII\\filter\\LandUseFilter_Allcats.txt",
             header = TRUE)

print(paste0("Found ", nrow(LUFilter), " pixels"))

arableLUFilter <- LUFilter %>%  filter(luCode == 3) 

print(paste0("Selected ", nrow(arableLUFilter), " arable pixels"))

head(arableLUFilter)

 for (p in 1:length(arableLUFilter$row)){     #loop rows (lat)
#  for (p in 1:3){     # test
 
  r <- arableLUFilter$row[p]
  c <- arableLUFilter$col[p]
    
 #   r <- 4
 #   c <- 130
    
    # Set control and storage variables  
    skipLoop <- FALSE # flag to get out of loops if NAs or -9999 is found
    thisLat <- NULL
    thisLong <- NULL
    
    Out <- data.frame(year = year, doy = doy) # add time dimensions 
     
    # Console info
    runTime <- round(Sys.time() - timeStart, digits = 2)
    print(paste("In row:", r," col:", c,
                " Duration in sec: ",
                runTime, sep = " ", collapse = NULL))
    timeStart <- Sys.time()
    
    # Loop factors
    for (f in 1:length(factorName)) {            
      if (skipLoop == TRUE) {next}  
      out.fact <- NULL # clear outputs of factors appended
      out.var <- NULL  # output of a variables with years appended
      
      #loop years 
      for (y in 1:length(firstYear)) {              
        if (skipLoop == TRUE) {next}  
        var <- NULL    # vector to hold factor variables (appended through years)
        
        # reads each netCDF file
        
        fileName <- paste(factorName[f],"_",firstYear[y],"-",firstYear[y]+9,".nc", sep = "", collapse = NULL)
                       
        # Console info
        print(paste("Running: ",fileName, sep = " ", collapse = NULL))
        
        netc <- nc_open(paste0(climDir,"//",fileName))
        
        time.dim <- length(ncvar_get(netc, "time"))
        thisLat <- netc$dim$latitude$vals[r]
        thisLong <-  netc$dim$longitude$vals[c] 
        
        # Do stuff with open netCDF file
        myVar <- NULL
        myVar <- ncvar_get(netc, start=c(c,r,1), count=c(1,1,-1))   
        # note: column is "X" and row is "Y" and time is "T" (see R netCDF doc)
        
        # Handle different NAs types used by NIWA
        if (is.na(myVar[1]) | myVar[1] == -9999) {
          
          # Console info
          print(paste("Found NA at: ",fileName,"row ",
                      r, "col: ",c," (skipping this grid-cell)",
                      sep = " ", collapse = NULL))         
          
          nc_close(base.nc)
          
          skipLoop <- TRUE # flag to skip the loop if missing data is found
          
          next
          
        } else {                  
          # Get data for all years in that grid-cell       
          #        print(myVar[1:10])
          
          # select variable and correct units
          switch(factorName[f],
                 rain={myVar = round(myVar, digits = 2)},
                 maxt={myVar = round(kToC(myVar), digits = 2)},              
                 mint={myVar = round(kToC(myVar), digits = 2)},
                 radn={myVar = round(WattToMJ(myVar), digits = 2)},
                      {print(paste("Unispecified factor:",
                      factorName[f], ", in year:", firstYear[y]))} 
                ) # end switch

  # appends variables daily in one vector         
               for (t in 1:time.dim) {     # loops through days t     
               var <- c(var, myVar[t])    
               }          

        } # end if not NAs

out.fact <- c(out.fact, var)  # appends all years for each variable

nc_close(base.nc)

      } # end year loop

# Appends data for all factors in one dataframe for each given grid-cell
# FIXME: Can this be in next loop?
if(skipLoop == FALSE) Out <- data.frame(Out, out.fact)

    } # end factors loop

# Prepare heading variables and co2

if(skipLoop == FALSE) { 
    
# Include the CO2 column for that scenario
  
# Option 1  (calculate CO2 from empirical fit)
# Out = data.frame(Out, co2 = round(co2Year(Out$year), digits = 1))
# Option 2: Read CO2 in txt file using year as lookup key
  Out <- merge(Out,co2PerRCP,by="year")  
  Out[,"co2"] <- round(Out[,"co2"], digits=2)
  
  # Define column names  <------ Attention here (FIXME: Should be automated)
  colnames(Out) <- c("year","day","maxt","mint","radn","rain", "co2")
  
  # Apsim header variables
  # FIXME: Should split climates by .met to avoid pooled averages?
  tav <- (colMeans(Out["maxt"]) + colMeans(Out["mint"]))*0.5
  jul <- subset(Out, doy > 182 & doy <212)
  dec <- subset(Out, doy > 335 & doy <364)
  tav.jul <- (colMeans(jul["maxt"]) + colMeans(jul["mint"]))*0.5 
  tav.dec <- (colMeans(dec["maxt"]) + colMeans(dec["mint"]))*0.5 
  amp <- tav.dec - tav.jul
   
  fileName <- paste0(r,"_",c,".met")
    
  # Construct heading
  site <- paste("NZ ! row=",r,"and col=", c)
  
  # These come from netCDF directly (ones that are used indeed)
  lat <- thisLat   # reading value from netCDF directly
  long <- thisLong
  
  # These values are for checking purposes only (Check in console)
  latCalc <- -34.375-0.05*(r-1) 
  longCalc <- 166.425+0.05*(c-1)
  print(paste0("Lat_Long in netCF (used): ", lat,"_",long,
               " and calculated: ", latCalc,"_",longCalc))
  
  # Do the heading (FIXME: Can we do it in a more tidy way?)
  writeLines(paste("[weather.met.weather]\nSite =", 
                   site,"! Climate from: ",climDir, 
                   "\nLatitude =", 
                   round(lat, digits = 2),
                   "(DECIMALDEGREES)\nLongitude = ",
                   round(long, digits = 2),
                   "(DECIMALDEGREES)\ntav= ", 
                   round(tav, digits = 2),
                   "(oC) ! annual  average  ambient  temperature\namp= ",
                   round(amp, digits = 2), 
                   "(oC) ! annual  amplitude in  mean  monthly  temperature\nyear  day  maxt  mint radn rain co2\n()  ()  (oC) (oC)  (MJ/m2/d) (mm) (ppm)",
                   sep = "",
                   collapse = NULL),
             paste0(outFolder,fileName))      
  
  # Print data in asc file
  write.table(Out, file = paste(outFolder,
                                fileName, sep = "", 
                                collapse = NULL),
              append = TRUE, 
              quote = FALSE,
              sep = " ", col.names = FALSE, 
              qmethod = "double", row.names = F)
  
    } # end if statement for missing value

#  }  # ends col loop

} # ends row loop 
```



```{r PrintRunDuration}
print(round(Sys.time()-timeStart), digits=1)
```





